---
title: "Gelman-Hill_Chap6"
author: "Sigurd Sørensen"
date: "9/28/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load packages}
pacman::p_load(tidyverse,haven)
#Haven has a function for reading .dta files
```

# Exercise 1
Poisson regression: the folder risky.behavior contains data from a randomized
trial targeting couples at high risk of HIV infection. The intervention provided
counseling sessions regarding practices that could reduce their likelihood
of contracting HIV. Couples were randomized either to a control group, a group
in which just the woman participated, or a group in which both members of
the couple participated. One of the outcomes examined after three months was
“number of unprotected sex acts.”

```{r}
risky_df <- read_dta("C:/Users/sigur/Desktop/github_methods_3/Gelman-Hill_Exercises/Data/risky_behaviors.dta")
```

## A
Predict fupacts(frequency of unprotected sex) following intervention. 

### Model Poisson
```{r}
risky_df<- risky_df %>% 
  mutate(fupacts_int = as.integer(fupacts))

m1 <- glm(fupacts_int ~ sex + couples + women_alone + bs_hiv, data = risky_df, family = poisson(link ="log"))
summary(m1)
exp(coef(m1))

```
### Test for overdispersion
Right now the Residual deviance is 12584 on 429 degree of freedom. Rule of thump
is that ratio between residual deviance and degree of freedom should be 1. But in
this case it is 30 indicating serious overdispersion. 

1) Here is a package for testing it. 
```{r}
pacman::p_load(AER)
dispersiontest(m1)
```
Now it indicates taht dispersion are even closer to 42.

2) Here is another package:

```{r}
pacman::p_load(DHARMa)
sim_m1 <- DHARMa::simulateResiduals(m1, refit = T)
testOverdispersion(sim_m1)
```
```{r}
plotSimulatedResiduals(sim_m1)
```
### Fix overdispersion
The quasi-families augment the normal families by adding a dispersion parameter. 
In other words, while for Poisson data Y¯ = s^2_Y, the quasi-Poisson allows
for Y¯ = τ · s^2_Y, and estimates the overdispersion
parameter τ (or underdispersion, if τ < 1).

https://biometry.github.io/APES/LectureNotes/2016-JAGS/Overdispersion/OverdispersionJAGS.pdf

```{r}
m2 <- glm(fupacts_int ~ sex + couples + women_alone + bs_hiv, data = risky_df, family = quasipoisson(link ="log"))
summary(m2)
```
It added a dispersion parameter similar to that calculated by the AER package. 

Once accounting for overdispersion sex and couples are no longer significant.
So not accounting for overdispersion could result in inflated p-values. 

## B 
Include bupacts as a predictor (unprotected sex prior to intervention)

We also see indication of overdispersion so we do quasipoisson regression. 
```{r}
m3 <- glm(fupacts_int ~ sex + couples + women_alone + bs_hiv + bupacts, data = risky_df, family = quasipoisson(link ="log"))
summary(m3)
```
## C
Interpret the coefficients of our overdispered quasipoisson regression model3.
```{r}
#Reversing from the log scale to get the scale back to the normal scale.
exp(coef(m3))
```
Intervention where both partners partcipate will increase the frequency of
unprotected sex by 0.66.

While the intervention of women_alone only increased it by 0.5136

MIGHT BE THE OTHER WAY AROUND SO THAT 0= Treatment and 1= No-treatment (would make sense)


## D
These data include responses from both men and women from the participating
couples. Does this give you any concern with regard to our modeling
assumptions?

Maybe I don't know...?!? 




# Exercise 2
Multinomial logit: using the individual-level survey data from the 2000 National
Election Study (data in folder nes), predict party identification (which is on a
five-point scale) using ideology and demographics with an ordered multinomial
logit model.




