---
title: "practical_exercise_5, Methods 3, 2021, autumn semester"
author: "Sigurd Fyhn Sørensen"
date: "13-10-2021"
output:
  html_document:
    df_print: paged
---

<style type="text/css">
  body{
  font-size: 14pt;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Exercises and objectives
The objectives of the exercises of this assignment are based on: https://doi.org/10.1016/j.concog.2019.03.007  
  
4) Download and organise the data from experiment 1  
5) Use log-likelihood ratio tests to evaluate logistic regression models  
6) Test linear hypotheses  
7) Estimate psychometric functions for the Perceptual Awareness Scale and evaluate them  

REMEMBER: In your report, make sure to include code that can reproduce the answers requested in the exercises below (__MAKE A KNITTED VERSION__)  
REMEMBER: This is part 2 of Assignment 2 and will be part of your final portfolio


# EXERCISE 4 - Download and organise the data from experiment 1
```{r packages, warning=FALSE}
pacman::p_load(tidyverse, lmerTest, lme4, stargazer, carret, readbulk, TinyTex, rstanarm, reabulk)
```


Go to https://osf.io/ecxsj/files/ and download the files associated with Experiment 1 (there should be 29).  
The data is associated with Experiment 1 of the article at the following DOI https://doi.org/10.1016/j.concog.2019.03.007  
  
1) Put the data from all subjects into a single data frame - note that some of the subjects do not have the _seed_ variable. For these subjects, add this variable and make in _NA_ for all observations. (The _seed_ variable will not be part of the analysis and is not an experimental variable)  
```{r, echo = FALSE, message = FALSE}
df_exp1 <- readbulk::read_bulk("experiment 1", extension = ".csv") 
set.seed(3215)
```
I identified some na's in the seed variable. 
```{r}
sum(is.na(df_exp1$seed))
```
precisely 2646. So no need to change anything as the NA's has aleady been added to the seed variable. 

    i. Factorise the variables that need factorising  

```{r preprocess}
df_exp1 <- df_exp1 %>% 
  mutate(right_answer = if_else(target.type == "odd" & obj.resp == "o" | target.type == "even" & obj.resp == "e",1,0)) %>% 
  mutate(right_answer = as.numeric(right_answer))


df_exp1 <- df_exp1 %>% 
  mutate(right_answer = as.numeric(right_answer)) %>% 
  mutate(subject = as.factor(subject)) %>% 
  mutate(task = as.factor(task)) %>% 
  mutate(target.type = as.factor (target.type)) %>% 
  mutate(trial.type = as.factor(trial.type)) %>% 
  mutate(pas = as.factor(pas))
```


    ii. Remove the practice trials from the dataset (see the _trial.type_ variable)
```{r}
levels(df_exp1$trial.type)
df_exp1 <- df_exp1 %>% 
  filter(trial.type == "experiment")

```

    iii. Create a _correct_ variable  
This has been done in the _r preprocess_ chunk. 

    iv. Describe how the _target.contrast_ and _target.frames_ variables differ compared to the data from part 1 of this assignment.  
```{r, echo = FALSE}
summary(df_exp1$target.contrast)
```
In experiment 2 (part 1 of this assignment) **target.frames** = 3 in all trials.
In experiment 1 (part 2 of this assignment) **target.frame** = (1:6) across trials.

In experiment 2 (part 1 of this assignment) target.contrast:
  Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
0.01000 0.05683 0.06329 0.08771 0.09392 1.00000 


In experiment 1 (part 2 of this assignment) target.contrast: 
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
    0.1     0.1     0.1     0.1     0.1     0.1 


# EXERCISE 5 - Use log-likelihood ratio tests to evaluate logistic regression models

1) Do logistic regression - _correct_ as the dependent variable and _target.frames_ as the independent variable. (Make sure that you understand what _target.frames_ encode). Create two models - a pooled model and a partial-pooling model. The partial-pooling model should include a subject-specific intercept.
```{r}
m5.1.1 <- glmer(right_answer ~ target.frames + (1|subject)+ (1|trial), data = df_exp1, family = binomial(link = "logit"))

m5.1.2 <- glm(right_answer ~ target.frames, data = df_exp1, family = binomial(link = "logit"))
```

    i. the likelihood-function for logistic regression is: $L(p)={\displaystyle\prod_{i=1}^Np^{y_i}(1-p)^{(1-y_i)}}$ (Remember the probability mass function for the Bernoulli Distribution). Create a function that calculates the likelihood. 
    
```{r}
likelihood_function <- function(model, y){
  p = fitted.values(model)
  temp_vec = p^y*(1-p)^(1-y)
  return(prod(temp_vec))
}
```
    
    ii. the log-likelihood-function for logistic regression is: $l(p) = {\displaystyle\sum_{i=1}^N}[y_i\ln{p}+(1-y_i)\ln{(1-p)}$. Create a function that calculates the log-likelihood  
```{r}
log_likelihood_function <- function(model, y){
  p = fitted.values(model)
  temp_vec = y * log(p) + (1-y) * log(1-p)
  return(sum(temp_vec))
}
```

    iii. apply both functions to the pooling model you just created. Make sure that the log-likelihood matches what is returned from the _logLik_ function for the pooled model. Does the likelihood-function return a value that is surprising? Why is the log-likelihood preferable when working with computers with limited precision?  
```{r}
likelihood_function(m5.1.2, df_exp1$right_answer)
log_likelihood_function(m5.1.2, df_exp1$right_answer)
```
I wouldn't say it is surprising that the likelihood function returns 0. If one iteration of i
in the product chain $L(p)={\displaystyle\prod_{i=1}^Np^{y_i}(1-p)^{(1-y_i)}}$ would be 0 the final product would result in 0. 

The argument for preferring log-likelihood is much similair to the argument of why likelihood returns 0.
Taking the log not only simplifies the subsequent mathematical analysis, but it also helps numerically because the product of a large number of small probabilities can easily be lesser than the numerical precision of the computer, and this is resolved by computing instead the sum of the log probabilities.

Log-likelihood is also preferential in regards of machine learning and optimization. 
It is less computational expensive and complicated to maximize a sum compared to a product.
Furthermore, we can make use of ln(ab)=ln(a)+ln(b). Now we can maximize simply by taking derivatives and setting equal to 0.

    iv. now show that the log-likelihood is a little off when applied to the partial pooling model - (the likelihood function is different for the multilevel function - see section 2.1 of https://www.researchgate.net/profile/Douglas-Bates/publication/2753537_Computational_Methods_for_Multilevel_Modelling/links/00b4953b4108d73427000000/Computational-Methods-for-Multilevel-Modelling.pdf if you are interested)  

```{r}
likelihood_function(m5.1.1, df_exp1$right_answer)
log_likelihood_function(m5.1.1, df_exp1$right_answer)

logLik(m5.1.1)
```
You can see the difference between -10563.5 (our function estimate) and another 
build in function logLik = -10622.03. 

2) Use log-likelihood ratio tests to argue for the addition of predictor variables, start from the null model, `glm(correct ~ 1, 'binomial', data)`, then add subject-level intercepts, then add a group-level effect of _target.frames_ and finally add subject-level slopes for _target.frames_. Also assess whether or not a correlation between the subject-level slopes and the subject-level intercepts should be included.
```{r}
m5.2.1 <- glm(right_answer ~ 1, family = binomial(link = "logit"), data = df_exp1)
m5.2.2 <- glm(right_answer ~ target.frames , family = binomial(link= "logit"), data = df_exp1)
m5.2.3 <- glmer(right_answer ~ target.frames + (1|subject), family = binomial(link = "logit"), data = df_exp1)

m5.2.4 <- glmer(right_answer ~ target.frames  + (1+ target.frames|subject), family = binomial(link = "logit"), data = df_exp1)
summary(m5.2.4)
```
 Following the guideline of the lme4 creators approach to selecting correlated or uncorrelated 
random effects in your model. https://cran.r-project.org/web/packages/lme4/vignettes/lmer.pdf (page. 7)
The correlation between target.frames and intercept is fairly high (-0.23) therefore
this needs to be specified in our model by specifying (x|y). This isn't assessed by log-likelihood
but rather correlation test between slope estimates and intercept. 

```{r, warning= FALSE}
anova( m5.2.4, m5.2.3, m5.2.1, m5.2.3)
```


A model calculating the log-likelihoood ratio test. Since all our models are nested
I've conditioned the order of variables following log-likelihood. If models weren't
nested I would need to order by the amount of columns in the design matrix (X+Z).

```{r, warning= FALSE}
log_like_ratio_test <- function(model1, model2){
  ratio = ifelse( logLik(model2) <= logLik(model1) , -2*(logLik(model2) - logLik(model1)) , (-2*(logLik(model1) - logLik(model2)))) #log-like-ratio
  
  df_val = abs(df.residual(model1) - df.residual(model2)) #calculate the degree of freedom
  p_val = pchisq(ratio, df = df_val, lower.tail = FALSE) #chi-square test
  name = deparse(substitute(c(model1, model2)))
  return(c(model_comp = name, log_lik_ratio = ratio, p_value = p_val))
}


log_like_ratio_test(m5.2.4, m5.2.1)
log_like_ratio_test(m5.2.4, m5.2.2)
log_like_ratio_test(m5.2.4, m5.2.3)
```


```{r}
#With package
pacman::p_load(lmtest)
lrtest(m5.2.3 , m5.2.4)
```
As shown, my functions output is similar to that of the packages lmtest::lrtest().




    i. write a short methods section and a results section where you indicate which model you chose and the statistics relevant for that choice. Include a plot of the estimated group-level function with `xlim=c(0, 8)` that includes the estimated subject-specific functions.
    
**logLik:**We performed a log-likelihood ratio test on our nested models with the most complex model as reference level. With a significance level of 0.05 we will reject the null-hypothesis 
on all comparisons. This supports the choice of selecting the most complex ie. model4.


**AIC:**While log-likelihood ratio does account for model complexity by chi-square distribution 
varying depending on Df. Aikai-Information-Criteria (AIC) has shown to be a better measure for
penalizing model complexity. the single level intercept model has the lowest AIC score
and will therefore be the preferred model following AIC. 


```{r}
newdata1 <- data.frame(target.frames = rep(seq(0,8,0.01),29) , subject = as.character(rep(seq(1:29),801))) %>% 
  mutate(subject = as.factor(subject))
#predict values for each subject target.frames 0:8
newdata1$yhat <- predict(m5.2.4, newdata = newdata1, type = "response")
```


```{r}
ggplot(newdata1, aes(x = target.frames, y = yhat, color = subject)) + geom_line() + 
  ylim(0,1) + labs(title = "group-level function with individual slope and intercept", y = "% of right answer = 1")
```
While this plot is well suited for looking at the variance between subjects slope and intercept
it is difficult to distinguish subjects from one another. If we facet by subject it will
allow individual inspection. 
```{r}
ggplot(newdata1, aes(x = target.frames, y = yhat)) + geom_line(aes(colour = subject)) + labs(y = "predicted propability of right answer") +  facet_wrap(~subject) + xlim(0,8) 
  
```
While graphs are nice numbers can also be very informative in illustrating the spread in our
beta parameters b0 and b1. 
```{r}
#get coef 
coef_m5.2.4 <- coef(m5.2.4)
slope <- invlogit(coef_m5.2.4$subject[,2])
intercept <- invlogit(coef_m5.2.4$subject[,1])
cbind(intercept = summary(intercept), slope = summary(slope))

```
Supported by the plots and statistical summary we can can conclude that there isn't
that much spread in the individual variance at subject level. 


**Funny Plot** This plot concatenates all subjects individual curve. But sadly 
they overlap each other a lot and make it hard to distinguish between them. If you look
closely you will be able to see 29 lines. But the 2 former plots are clearly better. :P 
```{r}
pacman::p_load("sjPlot")

(pp <- plot_model(m5.2.4 ,type="pred",
       terms=c("target.frames","subject"),pred.type="re"))
```



    ii. also include in the results section whether the fit didn't look good for any of the subjects. If so, identify those subjects in the report, and judge (no statistical test) whether their performance (accuracy) differed from that of the other subjects. Was their performance better than chance? (Use a statistical test this time) (50 %)  
```{r}
pacman::p_load(caret)
```

Generating an idea of accuracy in the entire data-set. 
```{r}
df_exp1<- df_exp1 %>% 
  mutate(fitted_m.5.2.4 = fitted(m5.2.4)) %>% 
  mutate(bin_fit_m5.2.4 = if_else(fitted_m.5.2.4 >= 0.5, 1, 0))

confusionMatrix(data = as.factor(df_exp1$bin_fit_m5.2.4), reference = as.factor(df_exp1$right_answer))
```

We illustrate differences in subject accuracy
```{r}
df_exp_temp <- df_exp1 %>% 
  group_by(subject) %>% 
  summarise(Percentage_correct_per_subject = sum(bin_fit_m5.2.4 == right_answer)/length(right_answer)*100)

ggplot(df_exp_temp, aes(x = subject, y = Percentage_correct_per_subject, fill = subject)) + geom_col()

```
Based on a visual inspection of predicted correct answers we can see that subject 24
has the worst predicted accuracy score. We will therefore perform as one-sample test 
to see if he performs better than chance. 
```{r}
df_exp1_fil24 <- df_exp1 %>% 
  filter(subject == 24)

#t-test on collected data
t.test(as.numeric(df_exp1_fil24$right_answer), mu = 0.5, alternative = "greater")
#t-test on predicted values. 
t.test(as.numeric(df_exp1_fil24$bin_fit_m5.2.4), mu = 0.5, alternative = "greater")
```
**T.test Summary**

*collected data*
The accuracy for subject 23 were above chance (50%),
(t(863) = 26.944, p = 2.2e^16)

*predicted data*
The predicted accuracy for subject 23 were above chance (50%),
(t(863) = 26.275, p = 2.2e^16)


3) Now add _pas_ to the group-level effects - if a log-likelihood ratio test justifies this, also add the interaction between _pas_ and _target.frames_ and check whether a log-likelihood ratio test justifies this  

    i. if your model doesn't converge, try a different optimizer  
```{r}
m5.3.1 <- glmer(right_answer ~ pas*target.frames + (target.frames||subject), data = df_exp1, family = binomial(link = "logit"), control = glmerControl(optimizer="bobyqa"))



log_like_ratio_test(m5.3.1 , m5.2.4)
```
Our log-likelihood-ratio test support the choice of the more complex model. (m5.3.1) 
including the interaction between pas:target.frames and pas as a fixed effect. 

    ii. plot the estimated group-level functions over `xlim=c(0, 8)` for each of the four PAS-ratings - add this plot to your report (see: 5.2.i) and add a description of your chosen model. Describe how _pas_ affects accuracy together with target duration if at all. Also comment on the estimated functions' behaviour at target.frame=0 - is that behaviour reasonable?  
```{r}
df_exp1 <- df_exp1 %>% 
  mutate(pred_interaction_model = fitted.values(m5.3.1))


interactions::interact_plot(model = m5.3.1 , pred = "target.frames", modx = "pas")
```



While this plot is nice for showing the general interaction between pas:target.frames
it does not account for our random intercept and random slop of target.frames.
We will therefore continue with another plot that can illustrate these individual differences. 

```{r}
ggplot(df_exp1, aes(x = target.frames, y = pred_interaction_model, col = pas )) + geom_point() + facet_wrap(~subject) + xlim(0,8) + ylim(0,1)

```
As shown target.frames has individual 
# EXERCISE 6 - Test linear hypotheses

In this section we are going to test different hypotheses. We assume that we have already proved that more objective evidence (longer duration of stimuli) is sufficient to increase accuracy in and of itself and that more subjective evidence (higher PAS ratings) is also sufficient to increase accuracy in and of itself.  
We want to test a hypothesis for each of the three neighbouring differences in PAS, i.e. the difference between 2 and 1, the difference between 3 and 2 and the difference between 4 and 3. More specifically, we want to test the hypothesis that accuracy increases faster with objective evidence if subjective evidence is higher at the same time, i.e. we want to test for an interaction.  

1) Fit a model based on the following formula: `correct ~ pas * target.frames + (target.frames | subject))`
    i. First, use `summary` (yes, you are allowed to!) to argue that accuracy increases faster with objective evidence for PAS 2 than for PAS 1. 

```{r, include = FALSE}
summary(m5.3.1)
```

```{r}
coef_OG_scale <-  data.frame(intercept = rep(NA,29), pas2 = rep(NA,29), pas3 = rep(NA,29), pas4 = rep(NA,29), target.frames = rep(NA,29), pas2_target.frames = rep(NA,29), pas3_target.frames = rep(NA,29), pas4_target.frames = rep(NA,29))

for (i in 1:length(coef(m5.3.1)$subject)){
  coef_OG_scale[,i] = invlogit(coef(m5.3.1)$subject[,i])
}
```


```{r, include = FALSE}
coef(m5.3.1)
coef_OG_scale

```
All beta values were transformed from log-odds to probabilities through an inverse logit transformation. All beta values will be reported on the probability scale for easier interpretation. '

**Intercepts**
Pas is coded as a factor and we will therefore gain an intercept for all 4 levels of pas.
*general baseline intercept*
$\beta_0 (intercept/pas1) = 0.4714, z = -2.064, p = 0.0390 $ 
*Individual baseline intercept per subject*
$\beta_0 (intercept/pas1): Median = 0.47, Mean = 0.4716, sd = 0.011$ 
Every subject has their own baseline/intercept probabillity of giving the right  answer when target.frames = 0 and pas = pas1.  

$\beta_1 (intercept/pas2) = 0.3604, z = -6.471, p = 9.77e-11$ added on top of of their individual baseline for pas1
$\beta_2 (intercept/pas3) = 0.3771, z = -3.653, p = 0.000259 ***$ added on top of of their individual baseline for pas1
$\beta_3 (intercept/pas4) = 0.5658, z = 1.065, p = 0.286944$ added on top of of their individual baseline for pas1

**Slopes**
We've modeled an interaction between a categorical variable(pas1-4) and a numeric (target.frames) 
with a second level effect of random slope 
We will therefore have a individual slope of target.frame for each subject.
**main effect**
$\beta_4 (target.frames) = Median = 0.5282, Mean = 0.5269, sd = 0.0234$
**interaction effect**
$\beta_5 (target.frames:pas2) = 0.6102, z = 13.006, p < 2e-16 $
$\beta_6 (target.frames:pas3) = 0.6764, z = 16.336, p < 2e-16 $
$\beta_7 (target.frames:pas4) = 0.6776, z = 10.990, p < 2e-16 $


Currently both genereal and individual intercept are below chance (50%). An explanation
can be found in the interpritation. Currently our intercept represent target.frames = 0 and
pas = pas1. Having been presented with 0 frames does not make much sense in terms of
the experiment. But we would expect a 50/50 chance to guess right if you had no
information to guess by. The best solution would be to standardize the data around target.type. 
Giving us a more intuitive interpritation.

Our model supports our assumption of an increase in subjective (pas-score) and objective (target.frames) evidence significantly increases the probability of the right answer.

Furthermore, as shown by our interaction effects increasing subjective evidence
(PAS1 -> PAS2/3/4) will significantly modulate objective evidence to increase the 
accuracy score even faster compared to PAS1. 

2) `summary` won't allow you to test whether accuracy increases faster with objective evidence for PAS 3 than for PAS 2 (unless you use `relevel`, which you are not allowed to in this exercise). Instead, we'll be using the function `glht` from the `multcomp` package
    i. To redo the test in 6.1.i, you can create a _contrast_ vector. This vector will have the length of the number of estimated group-level effects and any specific contrast you can think of can be specified using this. For redoing the test from 6.1.i, the code snippet below will do
```{r}
design <- model.matrix(m5.3.1)
head(design)
```

```{r}
pacman::p_load(multcomp)
contrast.vectors <- matrix(c(0, 0, 0, 0, -1, 1, 0, 0), nrow=1)
ghs <- glht(m5.3.1, contrast.vectors)
print(summary(ghs))
```
    
    ii. Now test the hypothesis that accuracy increases faster with objective evidence for PAS 3 than for PAS 2.

```{r}
#Contrast matrix


contrast_matrix_compar <- glht(m5.3.1, linfct = c( "pas2:target.frames = -1", "pas3:target.frames = 1"))
summary(contrast_matrix_compar) #not a 100% how this works. 

#Difference between pas2:target.frames pas3:target.frames
contrast.vector <- matrix(c(0, 0, 0, 0, 0, -1, 1, 0), nrow=1)
gh <- glht(m5.3.1, contrast.vector)
print(summary(gh))

```
objective evidence is also modulated more by PAS3 than PAS2 resulting in a faster growing 
accuracy by target.frames in PAS3 and than PAS2. 
    
    iii. Also test the hypothesis that accuracy increases faster with objective evidence for PAS 4 than for PAS 3
```{r}
#Difference between pas2:target.frames pas3:target.frames
contrast.vector2 <- matrix(c(0, 0, 0, 0, 0, 0, -1, 1), nrow=1)
gh2 <- glht(m5.3.1, contrast.vector2)
print(summary(gh2))
```
Moving from PAS3-PAS4 does not increase the effect of target.frames significantly. 

    
3) Finally, test that whether the difference between PAS 2 and 1 (tested in 6.1.i) is greater than the difference between PAS 4 and 3 (tested in 6.2.iii)
```{r}
# beta = 0.44821   std.error =  0.03446 pas1-pas2
# beta = 0.005258   std.error = 0.073618 pas3-pas4
confint_gh<- confint(gh)
confint_gh2 <- confint(gh2)

df_confint <- data.frame(beta = c(confint_gh$confint[1], confint_gh2$confint[1]), 
           upper = c(confint_gh$confint[2], confint_gh2$confint[2]), 
           lower = c(confint_gh$confint[3], confint_gh2$confint[3]),
           PAS1_2 = c(1,0),
           PAS3_4 = c(0,1))

ggplot(df_confint, aes(x = PAS1_2, y = beta)) + geom_point() + geom_errorbar(aes(x = PAS1_2, ymin = lower, ymax = upper), width = 0.2, color = "red", size = 1)
```
I can not do a t-test as calculating SE_pooled requires the SD. $SE_p = SD_p * sqrt(1/n_1 + 1/n_2)$.
We can calculate the 95% confidence intervals based on SE x 1.96 for lower and high boundaries. 
Our Confidence intervals does not overlap which would indicate that they are significantly different. 
THe difference of **PAS1-2** is larger than **PAS3-4**. 

# EXERCISE 7 - Estimate psychometric functions for the Perceptual Awareness Scale and evaluate them  

We saw in 5.3 that the estimated functions went below chance at a target duration of 0 frames (0 ms). This does not seem reasonable, so we will be trying a different approach for fitting here.  
We will fit the following function that results in a sigmoid, $f(x) = a + \frac {b - a} {1 + e^{\frac {c-x} {d}}}$  
It has four parameters: _a_, which can be interpreted as the minimum accuracy level, _b_, which can be interpreted as the maximum accuracy level, _c_, which can be interpreted as the so-called inflexion point, i.e. where the derivative of the sigmoid reaches its maximum and _d_, which can be interpreted as the steepness at the inflexion point. (When _d_ goes towards infinity, the slope goes towards a straight line, and when it goes towards 0, the slope goes towards a step function).  
  
We can define a function of a residual sum of squares as below

```{r}
RSS <- function(data, par)
{
    ## "dataset" should be a data.frame containing the variables x (target.frames)
    ## and y (correct)
    
    ## "par" are our four parameters (a numeric vector) 
    ## par[1]=a, par[2]=b, par[3]=c, par[4]=d
    a = par[1]
    b = par[2]
    c = par[3]
    d = par[4]
    x <- data$x
    y <- data$y
    y.hat <- a + ((b-a)/(1+(exp((c-x)/d))))
    RSS <- sum((y - y.hat)^2)
    return(RSS)
}
```

1) Now, we will fit the sigmoid for the four PAS ratings for Subject 7
    i. use the function `optim`. It returns a list that among other things contains the four estimated parameters. You should set the following arguments:  
    `par`: you can set _c_ and _d_ as 1. Find good choices for _a_ and _b_ yourself (and argue why they are appropriate)  
    `fn`: which function to minimise?  
    `data`: the data frame with _x_, _target.frames_, and _y_, _correct_ in it  
    `method`: 'L-BFGS-B'  
    `lower`: lower bounds for the four parameters, (the lowest value they can take), you can set _c_ and _d_ as `-Inf`. Find good choices for _a_ and _b_ yourself (and argue why they are appropriate)  
    `upper`: upper bounds for the four parameters, (the highest value they can take) can set _c_ and _d_ as `Inf`. Find good choices for _a_ and _b_ yourself (and argue why they are appropriate)   
```{r}
#Optim function with target.frames
data <- df_exp1 %>%
  dplyr::select(target.frames, right_answer) %>% 
  rename(x = target.frames, y = right_answer)


optim(par = c(0.5,1,1,1), fn = RSS, data = data, method = "L-BFGS-B", lower = c(0.5,0.5,-Inf,-Inf),
      upper = c(1,1,Inf,Inf))
```
```{r}
#Optim function with pas rating. 
data1 <- df_exp1 %>%
  filter(subject == 7) %>% #only use subject 7
  dplyr::select(pas, right_answer) %>% 
  mutate(pas = as.numeric(pas)) %>% #optim function requires it to be numeric. 
  rename(x = pas, y = right_answer)


optim_pas <- optim(par = c(0.5,1,1,1), fn = RSS, data = data1, method = "L-BFGS-B", lower = c(0.5,0.5,-Inf,-Inf),
      upper = c(1,1,Inf,Inf))
optim_pas$par
```
par = c(0.5, 1,1,1,) 
_a_ will start at 0.5 with a min = 0.5 and max = 1.
_b_ will start at 1 with min = 0.5 and max = 1.
The choice of min and max is based on the assignments wish to avoid values below chance.
So max and min values of y can span from 0.5:1. It is imaginable that such values could be possible 
so therefore they have been chosen as the lower and upper boundraies. 

Starting point is based on guesses of what the best scenario could be and worst scenario. 
ie. _a_ = 0.5 and _b_ = 1. 


Use the estimated parameters to compute y_hat and plot the function. 
```{r}
sigmoid <- function(a,b, c,d,x){ 
  y = a + ((b-a)/(1+(exp((c-x)/d))))
  return(y)
}

x_pas_optim <- seq(0, 4, 0.01)
y_pas_optim <- sigmoid(optim_pas$par[1],optim_pas$par[2],
                       optim_pas$par[3],optim_pas$par[4],x_pas_optim) 

```

```{r}
plot(data1$x, data1$y, pch = 16, xlab = "Pas Rating", ylab = "% right answer = 1")
lines(x_pas_optim, y_pas_optim) + title(main = "Sigmoid function for pas-ratings.", sub = "plot1")
```

    ii. Plot the fits for the PAS ratings on a single plot (for subject 7) `xlim=c(0, 8)`
```{r}
#Simple model to do illustrate pas ratings for subject 7
m7.1.1 <- glm(right_answer ~ as.numeric(pas), data = filter(df_exp1, subject == 7), family = binomial(link = "logit"))
x_pas <- seq(0, 4, 0.01)
y_pas <- predict(m7.1.1, list(pas = x_pas),type="response")
```

```{r}
#plots for subject 7
plot(as.numeric(df_exp1$pas), df_exp1$right_answer, pch = 16, xlab = "Pas Rating", ylab = "% right answer = 1", 
     xlim = c(0,8)) + title(main = "subject 7 glm function with no interaction", sub = "plot 2")

lines(x_pas, y_pas)
```
    
    iii. Create a similar plot for the PAS ratings on a single plot (for subject 7), but this time based on the model from 6.1 `xlim=c(0, 8)`   

```{r}
newdat <- data.frame(cbind('target.frames' = seq(0, 8, by = 0.001), 'pas' = rep(1:4), 'subject' = rep('7')))

newdat$subject <- as.factor(newdat$subject)
newdat$pas <- as.factor(newdat$pas)
newdat$target.frames <- as.numeric(newdat$target.frames)

newdat$yhat <- predict(m5.3.1, newdata = newdat, type = "response")
```

```{r}
ggplot(newdat, aes(x = target.frames, y = yhat, color = pas)) + geom_line() + labs(title = "Plot for subject 7, plot 4", y = "% of right_answer = 1")
```
    
    iv. Comment on the differences between the fits - mention some advantages and disadvantages of each way 
**Comparing plot1 (sigmoid) and plot2 (subject 7):**
**Plot 1** shows that as PAS approaches 4 the probability of guessing right approaches 100%. This is also illustrated in **plot 3** but also adds the detail of how PAS4 interacts with target.frames. While higher level of target.frames along with PAS4 rating will result in the best probability of guessing right. 

While **plot 1** gives us the best idea of how PAS rating effects probability of guessing correct. 
**plot 3** is more informative when it comes to illustrating the combination of target.frames and PAS. 
But as we will show in the next exercise there is a way to combine the information from both plots 
into 1 single plot. 

    
2) Finally, estimate the parameters for all subjects and each of their four PAS ratings. Then plot the estimated function at the group-level by taking the mean for each of the four parameters, _a_, _b_, _c_ and _d_ across subjects. A function should be estimated for each PAS-rating (it should look somewhat similar to Fig. 3 from the article:  https://doi.org/10.1016/j.concog.2019.03.007)

    i. compare with the figure you made in 5.3.ii and comment on the differences between the fits - mention some advantages and disadvantages of both.
```{r}
#Create a function with a loop for calculating a,b,c and d for every subject within every level of PAS. 

optim_for_indi <- function(){
  data_frame_parameters <- data.frame(subject = NA, pas = NA , a = NA, b = NA, c = NA, d = NA)
  
  for (i in 1:4){
    df_temp1 <- df_exp1 %>% 
      filter(pas == i)
    
    for (ii in 1:length(unique(df_exp1$subject))){
      data_temp <- df_temp1 %>% 
        filter(subject == ii) %>%
        dplyr::select(target.frames, right_answer, pas) %>% 
        rename(x = target.frames, y = right_answer)
        
      op_temp = optim(par = c(0.5,0.5,1,1), fn = RSS, data = data_temp, method = "L-BFGS-B", lower = c(0.5,0.5,-Inf,-Inf),
      upper = c(1,1,Inf,Inf))
      data_frame_parameters <- rbind(data_frame_parameters , c(ii,i,op_temp$par))
    
    }
  }
  return(data_frame_parameters)
}

```

```{r}
#Get all parameters for each subject in each level of pas
df_param <- optim_for_indi() %>% 
  na.omit()
df_param
#Compute the average value of a,b,c and d. 
df_param_avg <- df_param %>% 
  group_by(pas) %>%
  summarise(a = mean(a), b = mean(b), c = mean(c), d = mean(d))

df_param_avg
```
We currently have a problem of _a_ > _b_ in PAS1. This will result in a declining sigmoid function.
However, I can not specify logical statement in lower_boundaries ensuring _b_ > _a_. 

```{r}
#x-values
x_target.frames <- seq(0,8,0.01)


#setup data frame for x and y of different PAS score. 
df_x_y <- data.frame(x_target.frames)

#Function for estimating y_hat given our a,b,c and d from our optim function. 
y_hat_func <- function(){
for (i in 1:4){
  y_hat_temp <- sigmoid(df_param_avg$a[i], df_param_avg$b[i], df_param_avg
                        $c[i], df_param_avg$d[i], x_target.frames)
  df_x_y[,i+1] <- y_hat_temp
}
  
  df_x_y <- df_x_y %>% #Rename column names into something meaningful. 
    rename(x = x_target.frames, y_hat_pas1 = V2 , y_hat_pas2 = V3 , y_hat_pas3 = V4 , y_hat_pas4 = V5)
  return(df_x_y)
}

df_final_param <- y_hat_func()
```

```{r}
head(df_final_param)
```
```{r}
tail(df_final_param)
```
As X -> 8 our y_hat_pas1 estimates based on the sigmoid function defined in earlier chunks will 
decrease. This does not make sense what so ever but is due to _a_ and _b_ from our optim function. 


```{r}
df_final_param_long <- df_final_param %>% 
  pivot_longer(cols = c(y_hat_pas1,y_hat_pas2,y_hat_pas3,y_hat_pas4) , names_to = "pas_name", values_to = "y_hat_merged")

df_final_param_long
ggplot(df_final_param_long, aes(x, y_hat_merged, colour = pas_name)) +
  geom_line() + ylim(0,1) + labs(x = "target.frames", title = "averaged parameters across subject")
```



##### Extra stuff
Checking the difference between taking the mean() of every parameter c(a,b,c,d) and
running the optim on data only divided by PAS ratings. 

```{r}
#Make function

optim_pas_overall <- function(i){
  #prepare correct subset of  data frame.
  pas_temp <- df_exp1 %>% 
    filter(pas == i) %>% 
    dplyr::select(target.frames, right_answer) %>%
    rename(x = target.frames, y = right_answer)
  op_temp_pas = optim(par = c(0.5,1,1,1), fn = RSS, data = pas_temp, method = "L-BFGS-B", lower =  c(0.5,0.5,-Inf,-Inf), upper = c(1,1,Inf,Inf)) 
  
  # x and y values
  x_pas_optim <- seq(0, 8, 0.01)
  y_pas_optim <- sigmoid(op_temp_pas$par[1],op_temp_pas$par[2],
                       op_temp_pas$par[3],op_temp_pas$par[4],x_pas_optim) 
  return(y_pas_optim)
}



```

```{r}
#Generate y_hats with function for all pas ratings. 
data_frame_y_hats <- tibble(x = x_target.frames, y_hat_pas1 = optim_pas_overall(1)
                            , y_hat_pas2 = optim_pas_overall(2)
                            , y_hat_pas3 = optim_pas_overall(3)
                            , y_hat_pas4 = optim_pas_overall(4))
```


```{r}
#Pivot longer to merge into 1 columns with an identity column.
data_frame_y_hats_long <- data_frame_y_hats %>% 
  pivot_longer(cols = c(y_hat_pas1,y_hat_pas2,y_hat_pas3,y_hat_pas4) , names_to = "pas_name", values_to = "y_hat_merged")

#plot
ggplot(data_frame_y_hats_long, aes(x, y_hat_merged, colour = pas_name)) +
  geom_line() + ylim(0,1) + labs(x = "target.frames", title = "Parameters estimated grouped by PAS")
```
**Advantage**
If the relation between _% of guessing correct_ and _target.frames_ is in nature a sigmoid function.
ie. When target.frames goes from 0 -> _inf_ then would derivative first be fairly low
(difference between 0-1 isn't that great) while at some point we would hit the inflection point 
(point with highest derivative and effect of increasing target.frame by any unit.) And afterwards the 
derivative would lessen because increasing target.frames above a specific number would not 
result in any new information and therefore better accuracy. 

**Disadvantage/problem**
This plot/method alows us to set a minimum of our y-values. I've specified lower boundary 
of a = 0.5 (not allowing below chance). As this was our goal with using the optim function. Though there are several arguments against setting a minimum y-value. (It's cheating. :P )

The Optim function tries to minimise our RSS given our other parameters (a,b,c,d).
So if we really were to compare the model from optim with min = 0.5 and our interaction
model from exercise 5 we should look at diff in RSS. Specifying absurd high/low 
as _a_ & _b_ = 1 would give incredible accuracy by may have inflated RSS and poor
prediction accuracy of unseen data. 


